Titeln på avhandlingens språk

### 2023-02-07 body/part1/background.tex
fixing bookmarks;

@@

\medskip
Research in \glspl{NCS} has mostly focused on addressing the challenges imposed by the introduction of best-effort communication networks into traditional control systems.
Network delays (latency), losses, and \emph{jitter} --- referring to the variation in random network delays --- are at the forefront of these challenges.
They have partially been addressed through the clever placement of the controller (e.g.\ by employing edge computing)~\cite{sasaki2017layered,sasaki2016vehicle}, as well as through the development of control algorithms capable of accounting for the effects of the network~\cite{zhang2013network}.
\todo[inline]{TODO FINISH}

@@

\subsubsection{Related methodologies and tools}

The work we present in this thesis lies at the intersection of research on edge computing, \gls{WCA} applications, and \acl{CPS} such as \glspl{NCS}, each of these topics representing a vast body of research.
Of principal interest to us are works dealing with the experimental optimization of \gls{WCA} and \glspl{NCS} deployed on distributed systems such as edge computing.
Below we will summarize the key milestones in literature relating to this.

\paragraph{Experimental platforms for Edge Computing}

Research in edge computing is still in its early stages, although it has been steadily increasing in the last decade as researchers, businesses, and organizations recognize the benefits of bringing computing and storage capabilities closer to the source of data.
Standardization is however still a major unsolved challenge in this field, and thus most research hitherto has relied on ad-hoc methodologies and/or platforms.\todo{Citation}

Nonetheless, the last few years have seen the emergence of a number of small- to mid-scale platforms and testbeds for edge computing research.
Several of these are additionally driven by research interests in novel mobile networking technologies and \gls{MEC}, as well as in the virtualization of network functions.
Of particular interest to us are the
\begin{inlineenum}
    \item \acs{GENI}
    \item Chameleon
    \item \acs{COSMOS}
    \item \acs{POWDER}
    \item \acs{ARA}
    \item EdgeNet
%    \item Drexel Grid \gls{SDR}
\end{inlineenum} testbeds.

>>>>

\subsubsection{Experimental research in Edge Computing}

The last few years have seen the emergence of a number of small- to mid-scale platforms and testbeds for edge computing research.
Several of these are additionally driven by research interests in novel mobile networking technologies and \gls{MEC}, as well as in the virtualization of network functions.
Of particular interest to us are the
\begin{inlineenum}
    \item \acs{GENI}
    \item Chameleon
    \item \acs{COSMOS}
    \item \acs{POWDER}
    \item \acs{ARA}
    \item EdgeNet
%    \item Drexel Grid \gls{SDR}
\end{inlineenum} testbeds.

@@

\medskip
Research in \glspl{NCS} has mostly focused on addressing the challenges imposed by the introduction of best-effort communication networks into traditional control systems.
Network delays (latency), losses, and \emph{jitter} --- referring to the variation in random network delays --- are at the forefront of these challenges.
They have partially been addressed through the clever placement of the controller (e.g.\ by employing edge computing)~\cite{sasaki2017layered,sasaki2016vehicle}, as well as through the development of control algorithms capable of accounting for the effects of the network~\cite{zhang2013network}.

In terms of research relating to the benchmarking and optimization of these systems, works in literature have for the most part leveraged either theoretical models~\cite{zhang2019networked} or simulations (for example ~\cite{ma2019optimal}).
However, these approaches can only capture network behaviors at a coarse level, and thus a body of works has emerge in later years involving experimental approaches to \gls{NCS} benchmarking.
\cite{baumann2018evaluating} and~\cite{cuenca2019periodic} employ purpose-built physical testbeds for the study of these systems.
\cite{zoppi2020ncsbench} introduces \emph{NCSBench}, a platform for reproducible benchmarking of \gls{NCS} using a LEGO Mindstorms platform.

Finally, hybrid approaches, mixing testbed or prototype implementations and simulations or emulations of \gls{NCS} components also exist in literature.
These approaches tend to focus on a particular element of the system which is implemented using a prototype or testbed, and virtualize the rest of the system.
For instance, in~\cite{wang2020inverter} the authors are concerned with realistic effects from the network, and thus they have an emulated \gls{NCS} interact with a real network.
On the other hand, works such as~\cite{natale2004inverted} focus on the system being controlled and opt to instead emulate the network.

@@
